---
title: "Climate change in the Arctic: a poleward expansion of ticks and tick-borne diseases? - Immunological analyses"
author: "Amandine Gamble"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    theme: united
    toc: yes
editor_options:
  chunk_output_type: console
---

<style type="text/css">
  body{/* Normal */ font-size: 12px;}
  h1.title{font-size: 22px;}
  h1{/* Header 1 */ font-size: 18px;}
  h2{/* Header 2 */ font-size: 16px;}
  h3{ /* Header 2 */ font-size: 14px;}
  code.r{/* Code block */ font-size: 10px;}
  pre{/* Code block - determines code spacing between lines */ font-size: 10px;}
</style>

# Working space prepatation

```{r, warning = F, message = F}

# working directory to be adapted if needed
# setwd("")

# import data
sample_list = read.csv("data_clean/ELISA_arctic_borrelia_201510_sample_list.csv",sep = ";", header = T)
elisa_data = read.csv("data_clean/ELISA_arctic_borrelia_201510_ELISA_data.csv",sep = ";", header = T)

```

```{r}

knitr::opts_chunk$set(warning = F, message = F, fig.align = "center")

# list packages
packages = c("ggplot2", "patchwork", # plots
             "mixtools", "MASS", # mixture of distributions
             "lme4") # generalized linear mixed models

# install (if needed) and load packages
suppressMessages(invisible(lapply(
 packages,
 function(x) {
  if (!require(x, character.only = T)) {
   install.packages(x, dependencies = T)
   library(x, character.only = T)
  }
 }
)))

# plot settings
ggtheme = theme_bw() +
  theme(
    plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
    plot.margin = margin(5, 5, 5, 5),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.title.y.right = element_text(angle = 90),
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 8, color = "black"),
    axis.line = element_line(size = 0.25, color = "black"),
    axis.ticks = element_line(size = 0.25, color = "black"),
    strip.text = element_text(size = 8),
    strip.background = element_blank(),
    legend.margin = margin(2, 2, 2, 2),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key = element_rect(fill = NA, color = NA),
    legend.key.size = unit(0.4, "cm"),
    legend.background = element_rect(colour = "transparent", fill = ggplot2::alpha("white", 0))
  )

# color settings
col = c(neg = "#9ed99a", pos = "#8c34a9")

```

```{r}

elisa_data$plate = as.factor(elisa_data$plate)

# merge sample metadata and ELISA data
data_complete = merge(sample_list, elisa_data, by = c("sample_id"), all.x = T, all.y = T)
data_complete = subset(data_complete, !is.na(sample_id))

# check for missing samples
# View(data_complete[is.na(data_complete$OD_raw), ])
# View(data_complete[is.na(data_complete$sample_type), ])

data = aggregate(OD_adj ~ sample_id + plate + species + site + campaign, 
                 data = data_complete, 
                 FUN = "mean")

```

<br>
<br>

# 1. Interplate correction

<br>
<br>

## 1.1. Function for interplate regression

```{r}

get_reg=function(data,ref_plate){
  # replicates
  replicates=data[duplicated(data$sample_id) & !grepl("cont",data$sample_id),]$sample_id
  rep_data=data[is.element(data$sample_id,replicates),]
  
  rep_data = subset(rep_data, select = c("sample_id", "plate", "OD_adj"))
  
  regressions=as.data.frame(matrix(nrow=length(levels(rep_data$plate)),ncol=5))
  colnames(regressions)=c("plate","ref","intercept","slope","rsquared")
  
  OD_min = floor(min(rep_data$OD_adj)) 
  OD_max = ceiling(max(rep_data$OD_adj))
  
  for (i in 1:length(levels(rep_data$plate))){
    tryCatch({
      per_plate=merge(subset(rep_data,plate==levels(rep_data$plate)[i]),subset(rep_data,plate==ref_plate),by="sample_id")
      colnames(per_plate)=c("sample_id","plate","plate_OD","ref","ref_OD")
      reg=lm(plate_OD~ref_OD,data=per_plate)
      #plot(plate_OD~ref_OD,data=per_plate);text(plate_OD~ref_OD,labels=sample_id,cex=0.5,data=per_plate,pos=1)
      p=ggplot(data=per_plate,aes(x=ref_OD,y=plate_OD)) + 
        geom_point()+geom_smooth(method="lm",se=F,colour="black",linetype="dashed",size=0.5)+
        xlab(paste0("OD on Plate ",per_plate$ref))+ylab(paste0("OD on Plate ",per_plate$plate))+
        ggtitle("Interplate regression")+
        coord_fixed(ratio=1, xlim = c(OD_min, OD_max), ylim = c(OD_min, OD_max))+
        ggtheme
      print(p)
      
      regressions$plate[i]=levels(rep_data$plate)[i]
      regressions$ref[i]=ref_plate
      regressions$intercept[i]=reg$coefficients[1]
      regressions$slope[i]=reg$coefficients[2]
      regressions$rsquared[i]=summary(reg)$r.squared
    }, error=function(e){cat("ERROR : at plate",paste(levels(rep_data$plate)[i]),"\n")})
  }
  
  return(regressions)
}

#########################################################
# Calculate position and dispersion parameters

get_stat=function(data=NULL, measurevar, groupvars=NULL, na.rm=TRUE, conf.interval=.95, .drop=FALSE){
  
  for (i in 1:length(groupvars)){data[[groupvars[i]]]=as.factor(as.character(data[[groupvars[i]]]))}
  
  library(plyr)
  
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(n= length2(xx[[col]],na.rm=na.rm),
                     mean=round(mean(xx[[col]],na.rm=na.rm),digit=2),
                     sd=round(sd(xx[[col]], na.rm=na.rm),digit=2),
                     se=NA,
                     ci=NA,
                     min= min(xx[[col]], na.rm=na.rm),
                     q25=quantile(xx[[col]], na.rm=na.rm)[2][[1]],
                     median=median(xx[[col]], na.rm=na.rm),
                     q75=quantile(xx[[col]], na.rm=na.rm)[4][[1]],
                     max=max(xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  
  datac$se <- round(datac$sd / sqrt(datac$n),digit=2)
  
  ciMult <- qt(conf.interval/2 + .5, datac$n-1)
  datac$ci <- round(datac$se * ciMult,digit=2)
  
  return(datac)
}

```

<br>
<br>

## 1.2. Exploration of the plate effect on OD (before correction)

```{r,warning=FALSE,message=FALSE,fig.width=14*0.4,fig.height=7*0.4}

# select the interplate replicate samples
replicates = unique(data[duplicated(data$sample_id), ]$sample_id)

# null model explaining variations of OD accounting for a (random) sample effect
mod_no_plate_effect = lmer(OD_adj ~ (1 | sample_id), 
                           data = data[is.element(data$sample_id, replicates) == T, ])

# same model + accounting for a (fixed) plate effect
mod_plate_effect = lmer(OD_adj ~ plate + (1 | sample_id), 
                        data = data[is.element(data$sample_id, replicates) == T, ])
# model comparison
anova(mod_no_plate_effect, mod_plate_effect)

# violin plot
ggplot(data[is.element(data$sample_id, replicates) == T, ], aes(y = OD_adj, x = plate)) +
  geom_violin() + geom_point(aes(color = sample_id)) +
  xlab("Plate") + ylab("Adjusted OD (/plate controls)") +
  ggtitle("Before correction") +
  ggtheme + theme(legend.position = "none")

```

<br>
<br>

## 1.3. Interplate regressions and correction

```{r,warning=FALSE,message=FALSE,fig.width=7*0.4,fig.height=7*0.4}

# fit the regressions using th home-made function get_reg()
regressions=get_reg(data,"3")
regressions

# correct the OD based on the regressions
for (i in 1:nrow(regressions)){
  plate=data[data$plate==regressions$plate[i],]
  plate$ref=regressions$ref[i]
  plate$OD_cor=(-regressions$intercept[i]+plate$OD_adj)/regressions$slope[i]
  
  if (i==1){data_corrected=plate
  }else{data_corrected=rbind(data_corrected,plate)}
}
remove(i)

# formatting
data_corrected$plate=as.factor(data_corrected$plate)

```

<br>
<br>

```{r,warning=FALSE,message=FALSE,fig.width=14*0.4,fig.height=7*0.4}

# plot before correction
ggplot(subset(data_corrected,is.element(sample_id,replicates)),aes(x=as.numeric(plate),y=OD_adj,colour=sample_id))+
  geom_point()+geom_line()+
  xlab("Plate")+ylab("Adjusted OD (/plate controls)")+
  ggtitle("Before correction")+
  ggtheme

# plot after correction
ggplot(subset(data_corrected,is.element(sample_id,replicates)),aes(x=as.numeric(plate),y=OD_cor,colour=sample_id))+
  geom_point()+geom_line()+
  xlab("Plate")+ylab("Corrected OD (/interplate variations)")+
  ggtitle("After correction")+
  ggtheme

remove(i)

```

<br>
<br>

## 1.4. Exploration of the plate effect on corrected OD (after correction)

```{r,warning=FALSE,message=FALSE,fig.width=14*0.4,fig.height=7*0.4}

# model fitting and comparison (similarly as above but on corrected OD)
mod_sample_effect_cor=lmer(OD_cor~(1|sample_id),data=data_corrected[is.element(data_corrected$sample_id,replicates)==T,])
mod_plate_effect_cor=lmer(OD_cor~plate+(1|sample_id),data=data_corrected[is.element(data_corrected$sample_id,replicates)==T,])
anova(mod_sample_effect_cor,mod_plate_effect_cor)

# violin plot
ggplot(data_corrected[is.element(data_corrected$sample_id,replicates)==T,],aes(y=OD_cor,x=plate))+
  geom_violin()+geom_point()+
  xlab("Plate")+ylab("Corrected OD (/interplate variations)")+
  ggtitle("After correction")+
  ggtheme

# box plot
ggplot(data_corrected[is.element(data_corrected$sample_id,replicates)==T,],aes(y=OD_cor,x=plate))+
  geom_boxplot()+geom_point()+
  xlab("Plate")+ylab("Corrected OD (/interplate variations)")+
  ggtitle("After correction")+
  ggtheme

remove(mod_sample_effect,mod_plate_effect,mod_sample_effect_cor,mod_plate_effect_cor,plate,regressions,replicates)

```

<br>
<br>

# 2. ELISA cut-off 

Method described in <a href="https://doi.org/10.1111/1365-2435.12942" target="_blank">Garnier et al.,2017,Functional Ecology</a>

<br>
<br>

```{r,warning=FALSE,message=FALSE,fig.width=14*0.4,fig.height=7*0.4}

data_rissa_complete = aggregate(OD_cor~sample_id + site, subset(data_corrected, species == "Rissa tridactyla"), FUN = "mean")
data_uria_complete = aggregate(OD_cor~sample_id + site, subset(data_corrected, species == "Uria lomvia"), FUN = "mean")

data_rissa_ossian = subset(data_rissa_complete, site == "Ossian")
data_uria_ossian = subset(data_uria_complete,  site == "Ossian")

data = data_rissa_complete
# fits two normal distributions to the corrected OD values
mixnorm<-normalmixEM(subset(data,!is.na(OD_cor))$OD_cor,eps=1e-12,lambda=0.5,mu=c(0,1),sigma=c(0.1,1),k=2)
# fits one normal distribution to the same data as above
norm1<-fitdistr(subset(data,!is.na(OD_cor))$OD_cor,"normal")
# manually calculate AIC 
AIC1n<-2*2-2*norm1$loglik
AIC2n<-4*2-2*mixnorm$loglik
c(AIC1n,AIC2n)
if (AIC2n<AIC1n-2){print("Two normal distributions")}else{print("One unique normal distribution")}
cut95=mixnorm$mu[1]+2*mixnorm$sigma[1]
cut99=mixnorm$mu[1]+3*mixnorm$sigma[1]
c(cut95,cut99)

mixnorm_rissa_complete = mixnorm
cut95_rissa_complete=mixnorm_rissa_complete$mu[1]+2*mixnorm_rissa_complete$sigma[1]
cut99_rissa_complete=mixnorm_rissa_complete$mu[1]+3*mixnorm_rissa_complete$sigma[1]


# plot
ggplot(data, aes(x = OD_cor)) + 
  geom_histogram(aes(fill = site),
                 binwidth=.3,size=0.25,
                 colour="black")+
  stat_function(aes(colour="#8c34a9"),fun=dnorm,args=list(mean=mixnorm$mu[1],sd=mixnorm$sigma[1]),size=1)+
  stat_function(aes(colour="#9ed99a"),fun=dnorm,args=list(mean=mixnorm$mu[2],sd=mixnorm$sigma[2]),size=1)+
  geom_vline(aes(xintercept=cut99),linetype="dashed",size=0.5)+
  geom_vline(aes(xintercept=cut95),linetype="dashed",size=0.5)+
  ylab("Density")+xlab("Optical density (ELISA)")+
  ggtheme+
  scale_color_manual(values=c("#9ed99a","#8c34a9"),labels=c("ELISA-negative sample distribution","ELISA-positive sample distribution"))+
  scale_fill_manual(values = c("gray90", "grey50")) +
  theme(legend.justification=c(1,1),legend.position=c(1,1)) +
  ggtitle("Black-legged kittiwake")

x= seq(0, 6, 0.01)
pred_rissa_complete = data.frame(cbind(x = x, y1 = dnorm(x,mean=mixnorm_rissa_complete$mu[1],sd=mixnorm_rissa_complete$sigma[1])*2.5, y2 = dnorm(x,mean=mixnorm_rissa_complete$mu[2],sd=mixnorm_rissa_complete$sigma[2])*2.5))
p1 = ggplot() + 
  geom_histogram(data = data_rissa_complete, aes(x = OD_cor, fill = site),
                 binwidth=.1,size=0.25,
                 colour="black")+
  geom_line(data = pred_rissa_complete, aes(x=x, y=y1, colour="#8c34a9"), size = 0.75, alpha = 0.7) +
  geom_line(data = pred_rissa_complete, aes(x=x, y=y2, colour="#9ed99a"), size = 0.75, alpha = 0.7) +
  geom_vline(aes(xintercept=cut99_rissa_complete),linetype="dashed",size=0.5)+
  geom_vline(aes(xintercept=cut95_rissa_complete),linetype="dashed",size=0.5)+
  ylab("Number of samples")+xlab("Mean optical density (anti-Bbsl ELISA)")+
  ggtheme+
    scale_fill_manual(values = c("gray90", "grey50"), labels = c("Hornøya", "Ossian")) +
  scale_color_manual(values=c("#9ed99a","#8c34a9"),labels=c("ELISA-negative sample distribution","ELISA-positive sample distribution"))+
  theme(legend.justification=c(1,1),legend.position=c(1,1)) +
  ggtitle("Black-legged kittiwakes")+ ylim(c(0,25)) + xlim(c(0,6)) +
  annotate("text", x=0.2, y=25, label= "Negative", size = 2) +
  annotate("text", x=1.5, y=25, label= "Positive", size = 2) 


data = data_uria_complete 
# fits two normal distributions to the corrected OD values
mixnorm<-normalmixEM(subset(data,!is.na(OD_cor))$OD_cor,eps=1e-12,lambda=0.5,mu=c(0,1),sigma=c(0.1,1),k=2)
# fits one normal distribution to the same data as above
norm1<-fitdistr(subset(data,!is.na(OD_cor))$OD_cor,"normal")
# manually calculate AIC 
AIC1n<-2*2-2*norm1$loglik
AIC2n<-4*2-2*mixnorm$loglik
c(AIC1n,AIC2n)
if (AIC2n<AIC1n-2){print("Two normal distributions")}else{print("One unique normal distribution")}
cut95=mixnorm$mu[1]+2*mixnorm$sigma[1]
cut99=mixnorm$mu[1]+3*mixnorm$sigma[1]
c(cut95,cut99)


mixnorm_uria_complete = mixnorm
cut95_uria_complete=mixnorm_uria_complete$mu[1]+2*mixnorm_uria_complete$sigma[1]
cut99_uria_complete=mixnorm_uria_complete$mu[1]+3*mixnorm_uria_complete$sigma[1]

pred_uria_complete = data.frame(cbind(x = x, y1 = dnorm(x,mean=mixnorm_uria_complete$mu[1],sd=mixnorm_uria_complete$sigma[1])*5, y2 = dnorm(x,mean=mixnorm_uria_complete$mu[2],sd=mixnorm_uria_complete$sigma[2])*5))
p2 = ggplot() + 
  geom_histogram(data = data_uria_complete, aes(x = OD_cor, fill = site),
                 binwidth=.1,size=0.25,
                 colour="black")+
  geom_line(data = pred_uria_complete, aes(x=x, y=y1, colour="#8c34a9"), size = 0.75, alpha = 0.7) +
  geom_line(data = pred_uria_complete, aes(x=x, y=y2, colour="#9ed99a"), size = 0.75, alpha = 0.7) +
  geom_vline(aes(xintercept=cut99_uria_complete),linetype="dashed",size=0.5)+
  geom_vline(aes(xintercept=cut95_uria_complete),linetype="dashed",size=0.5)+
  ylab("Number of samples")+xlab("Mean optical density (anti-Bbsl ELISA)")+
  ggtheme+
    scale_fill_manual(values = c("gray90", "grey50"), labels = c("Hornøya", "Ossian")) +
  scale_color_manual(values=c("#9ed99a","#8c34a9"),labels=c("ELISA-negative sample distribution","ELISA-positive sample distribution"))+
  theme(legend.justification=c(1,1),legend.position=c(1,1)) +
  ggtitle("Brünnich guillemots")+ ylim(c(0,25)) + xlim(c(0,6)) +
  annotate("text", x=0.5, y=25, label= "Negative", size = 2) +
  annotate("text", x=2, y=25, label= "Positive", size = 2) 

# plot
ggplot(data, aes(x = OD_cor)) + 
  stat_function(aes(colour="#8c34a9"),fun=dnorm,args=list(mean=mixnorm$mu[1],sd=mixnorm$sigma[1]),size=0.5)+
  stat_function(aes(colour="#9ed99a"),fun=dnorm,args=list(mean=mixnorm$mu[2],sd=mixnorm$sigma[2]),size=0.5)+
  geom_histogram(aes(y=..density.., fill = site),
                 binwidth=.1,size=0.25,
                 colour="black")+
  geom_vline(aes(xintercept=cut99),linetype="dashed",size=0.5)+
  geom_vline(aes(xintercept=cut95),linetype="dashed",size=0.5)+
  ylab("Density")+xlab("Optical density (ELISA)")+
  ggtheme+
  scale_color_manual(values=c("#9ed99a","#8c34a9"),labels=c("ELISA-negative sample distribution","ELISA-positive sample distribution"))+
  scale_fill_manual(values = c("gray90", "grey50")) +
  theme(legend.justification=c(1,1),legend.position=c(1,1))

data = data_rissa_ossian
# fits two normal distributions to the corrected OD values
mixnorm<-normalmixEM(subset(data,!is.na(OD_cor))$OD_cor,eps=1e-12,lambda=0.5,mu=c(0,1),sigma=c(0.1,1),k=2)
# fits one normal distribution to the same data as above
norm1<-fitdistr(subset(data,!is.na(OD_cor))$OD_cor,"normal")
# manually calculate AIC 
AIC1n<-2*2-2*norm1$loglik
AIC2n<-4*2-2*mixnorm$loglik
c(AIC1n,AIC2n)
if (AIC2n<AIC1n-2){print("Two normal distributions")}else{print("One unique normal distribution")}
cut95=mixnorm$mu[1]+2*mixnorm$sigma[1]
cut99=mixnorm$mu[1]+3*mixnorm$sigma[1]
c(cut95,cut99)

# plot
ggplot(data, aes(x = OD_cor)) + 
  stat_function(aes(colour="#8c34a9"),fun=dnorm,args=list(mean=mixnorm$mu[1],sd=mixnorm$sigma[1]),size=0.5)+
  stat_function(aes(colour="#9ed99a"),fun=dnorm,args=list(mean=mixnorm$mu[2],sd=mixnorm$sigma[2]),size=0.5)+
  geom_histogram(aes(y=..density.., fill = site),
                 binwidth=.1,size=0.25,
                 colour="black")+
  geom_vline(aes(xintercept=cut99),linetype="dashed",size=0.5)+
  geom_vline(aes(xintercept=cut95),linetype="dashed",size=0.5)+
  ylab("Density")+xlab("Optical density (ELISA)")+
  ggtheme+
  scale_color_manual(values=c("#9ed99a","#8c34a9"),labels=c("ELISA-negative sample distribution","ELISA-positive sample distribution"))+
  scale_fill_manual(values = c("gray90", "grey50")) +
  theme(legend.justification=c(1,1),legend.position=c(1,1))

data = data_uria_ossian 
# fits two normal distributions to the corrected OD values
mixnorm<-normalmixEM(subset(data,!is.na(OD_cor))$OD_cor,eps=1e-12,lambda=0.5,mu=c(0,1),sigma=c(0.1,1),k=2)
# fits one normal distribution to the same data as above
norm1<-fitdistr(subset(data,!is.na(OD_cor))$OD_cor,"normal")
# manually calculate AIC 
AIC1n<-2*2-2*norm1$loglik
AIC2n<-4*2-2*mixnorm$loglik
c(AIC1n,AIC2n)
if (AIC2n<AIC1n-2){print("Two normal distributions")}else{print("One unique normal distribution")}
cut95=mixnorm$mu[1]+2*mixnorm$sigma[1]
cut99=mixnorm$mu[1]+3*mixnorm$sigma[1]
c(cut95,cut99)

# plot
ggplot(data, aes(x = OD_cor)) + 
  stat_function(aes(colour="#8c34a9"),fun=dnorm,args=list(mean=mixnorm$mu[1],sd=mixnorm$sigma[1]),size=0.5)+
  stat_function(aes(colour="#9ed99a"),fun=dnorm,args=list(mean=mixnorm$mu[2],sd=mixnorm$sigma[2]),size=0.5)+
  geom_histogram(aes(y=..density.., fill = site),
                 binwidth=.1,size=0.25,
                 colour="black")+
  geom_vline(aes(xintercept=cut99),linetype="dashed",size=0.5)+
  geom_vline(aes(xintercept=cut95),linetype="dashed",size=0.5)+
  ylab("Density")+xlab("Optical density (ELISA)")+
  ggtheme+
  scale_color_manual(values=c("#9ed99a","#8c34a9"),labels=c("ELISA-negative sample distribution","ELISA-positive sample distribution"))+
  scale_fill_manual(values = c("gray90", "grey50")) +
  theme(legend.justification=c(1,1),legend.position=c(1,1))

# fits two normal distributions to the corrected OD values
mixnorm<-normalmixEM(subset(data,!is.na(OD_cor))$OD_cor,eps=1e-12,lambda=0.5,mu=c(0,1),sigma=c(0.1,1),k=2)
# fits one normal distribution to the same data as above
norm1<-fitdistr(subset(data,!is.na(OD_cor))$OD_cor,"normal")
# manually calculate AIC 
AIC1n<-2*2-2*norm1$loglik
AIC2n<-4*2-2*mixnorm$loglik
c(AIC1n,AIC2n)
if (AIC2n<AIC1n-2){print("Two normal distributions")}else{print("One unique normal distribution")}
cut95=mixnorm$mu[1]+2*mixnorm$sigma[1]
cut99=mixnorm$mu[1]+3*mixnorm$sigma[1]
c(cut95,cut99)


mixnorm_rissa_ossian = mixnorm
cut95_rissa_ossian=mixnorm_rissa_ossian$mu[1]+2*mixnorm_rissa_ossian$sigma[1]
cut99_rissa_ossian=mixnorm_rissa_ossian$mu[1]+3*mixnorm_rissa_ossian$sigma[1]


mixnorm_uria_ossian = mixnorm
cut95_uria_ossian=mixnorm_uria_ossian$mu[1]+2*mixnorm_uria_ossian$sigma[1]
cut99_uria_ossian=mixnorm_uria_ossian$mu[1]+3*mixnorm_uria_ossian$sigma[1]

# plot
ggplot(data, aes(x = OD_cor)) + 
  stat_function(aes(colour="#8c34a9"),fun=dnorm,args=list(mean=mixnorm$mu[1],sd=mixnorm$sigma[1]),size=0.5)+
  stat_function(aes(colour="#9ed99a"),fun=dnorm,args=list(mean=mixnorm$mu[2],sd=mixnorm$sigma[2]),size=0.5)+
  geom_histogram(aes(y=..density.., fill = site),
                 binwidth=.1,size=0.25,
                 colour="black")+
  geom_vline(aes(xintercept=cut99),linetype="dashed",size=0.5)+
  geom_vline(aes(xintercept=cut95),linetype="dashed",size=0.5)+
  ylab("Density")+xlab("Optical density (ELISA)")+
  ggtheme+
  scale_color_manual(values=c("#9ed99a","#8c34a9"),labels=c("ELISA-negative sample distribution","ELISA-positive sample distribution"))+
  theme(legend.justification=c(1,1),legend.position=c(1,1))

#ggsave("export/elisa_OD_distribution.pdf",width=14,height=7,units="cm",dpi=300)

((p1) | (p2  + theme(axis.title.y = element_blank(), axis.text.y = element_blank())) ) + plot_layout(guides = "collect") & theme(legend.position = 'bottom', legend.direction="vertical")

ggsave("export/arctic_borrelia_ELISA_distribution.pdf",width=15,height=9,units="cm",dpi=300)

# save serological status
data_rissa_complete$sero_pos=NA
data_rissa_complete[!is.na(data_rissa_complete$OD_cor) & data_rissa_complete$OD_cor>cut99_rissa_complete,"sero_pos"]="positive"
data_rissa_complete[!is.na(data_rissa_complete$OD_cor) & data_rissa_complete$OD_cor<=cut99_rissa_complete,"sero_pos"]="doubtful"
data_rissa_complete[!is.na(data_rissa_complete$OD_cor) & data_rissa_complete$OD_cor<=cut95_rissa_complete,"sero_pos"]="negative"

data_uria_complete$sero_pos=NA
data_uria_complete[!is.na(data_uria_complete$OD_cor) & data_uria_complete$OD_cor>cut99_uria_complete,"sero_pos"]="positive"
data_uria_complete[!is.na(data_uria_complete$OD_cor) & data_uria_complete$OD_cor<=cut99_uria_complete,"sero_pos"]="doubtful"
data_uria_complete[!is.na(data_uria_complete$OD_cor) & data_uria_complete$OD_cor<=cut95_uria_complete,"sero_pos"]="negative"

write.table(data,"export/data.csv",quote=F,col.names=T,row.names=F,sep=";",dec=".")

remove(mixnorm,AIC2n,norm1,AIC1n)

length(unique(data_rissa_complete$sample_id))
length(unique(data_uria_complete$sample_id))
length(unique(data_rissa_ossian$sample_id))
length(unique(data_uria_ossian$sample_id))
sum(sample_list$species == "Uria lomvia")

```

<br>
<br>

# 3. Seroprevalences

<br>
<br>

```{r,warning=FALSE,message=FALSE}


get_prev=function(data=NULL, measurevar, groupvars=NULL, na.rm=TRUE, conf.interval=.95, .drop=FALSE){
  
  for (i in 1:length(groupvars)){data[[groupvars[i]]]=as.factor(as.character(data[[groupvars[i]]]))}
  
  library(plyr)
  library(PropCIs)
  
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  datac1 <- ddply(data, groupvars, .drop=.drop,
                  .fun = function(xx, col) {
                    c(n=length2(xx[[col]],na.rm=na.rm)
                    )
                  },
                  measurevar
  )
  
  datac2 <- ddply(subset(data,data[[measurevar]]==1 | data[[measurevar]]=="pos"), groupvars, .drop=.drop,
                  .fun = function(xx, col) {
                    c(pos=length2(xx[[col]],na.rm=na.rm)
                    )
                  },
                  measurevar
  )
  
  datac=merge(datac1,datac2)
  
  datac$prev <- round(datac$pos/datac$n,digit=2)
  
  for (i in 1:nrow(datac)){
    if (is.na(datac$prev[i])){datac$CImin[i]=NA;datac$CImax[i]=NA
    }else{
      datac$CImin[i] <- round(exactci(datac$pos[i],datac$n[i],conf.interval)$conf.int[1],digit=2)
      datac$CImax[i] <- round(exactci(datac$pos[i],datac$n[i],conf.interval)$conf.int[2],digit=2)      
    }
  }
  
  return(datac)
}

data_unique=subset(data_corrected,!duplicated(sample_id) & !grepl("cont",sample_id))

# overall seroprevalence
sum(data_unique$sero_pos==1)/sum(!is.na(data_unique$sero_pos))

# detailed seroprevalences
sum(data_rissa_complete$site == "Ossian" & data_rissa_complete$sero_pos == "positive")
sum(data_rissa_complete$site == "Ossian" & data_rissa_complete$sero_pos == "positive")/sum(data_rissa_complete$site == "Ossian")
sum(data_rissa_complete$site == "Ossian" & data_rissa_complete$sero_pos == "doubtful")

sum(data_uria_complete$site == "Ossian" & data_uria_complete$sero_pos == "positive")
sum(data_uria_complete$site == "Ossian" & data_uria_complete$sero_pos == "positive")/sum(data_uria_complete$site == "Ossian")
sum(data_uria_complete$site == "Ossian" & data_uria_complete$sero_pos == "doubtful")
sum(data_uria_complete$site == "Ossian")

```
